---
title: "Microarray Data - Axin2 Trp53 KO - GSEA analysis on DGE results - Plots"
author: "Hilmar Berger"
output: 
 html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    number_sections: true
    code_folding: hide
    
pdf_document:
    fig_caption: true
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
---

```{r, warning=FALSE, message=FALSE}
rm(list=ls())

#install.packages(c("ggplot2","knitr","scales","data.table","gridExtra","BiocManager") )
#BiocManager::install("fgsea","limma")


#suppressMessages(library(xlsx))
library(ggplot2)
library(knitr)
library(scales)
library(fgsea)
library(data.table)
library(gridExtra)

source("../../../../r_utils/fGSEA_plot_functions.R")

result_folder = "../../Results/Axin2_Trp53_KO"
set.seed(12345)
```


# Overview

This is data from a stomach corpus samples of mouse experiments (Axin2 Trp53 KO) hybridized as dual channel libraries to Agilent 074809 (SurePrint G3 Mouse GE v2 8x60K) micro arrays. 

For analysis, a preranked analysis using the R package fGSEA was used that should give similar results to preranked analysis in standard GSEA. 


```{r}
load("../../Results/Axin2_Trp53_KO/DGE_results.Rdata")
```

```{r}
output_folder = "../../Results/Axin2_Trp53_KO"
GSEA_datafile = file.path(output_folder,"fGSEA_Results.Rdata")
load(GSEA_datafile)
load(file.path(output_folder, "Geneset_collections.Rdata"))
```


# MSigDB gene sets

Since MSigDB gene sets use human gene symbols to map genes to pathways we translated mouse symbols to homologous human symbols using HomologeneDB from NCBI (build 68). 


```{r}
homologene_human_mouse = fread("../../Data/External/homologene_9606_10090.txt")
all_entrez_ids = data.frame(EntrezID=unique(all_results[[1]]$EntrezID))
all_symbols = data.frame(GeneSymbol=unique(all_results[[1]]$GeneSymbol))
a1 = merge(all_symbols, homologene_human_mouse[,c("entrez_1","symbol_1","symbol_2"), with=F], by.x="GeneSymbol", by.y="symbol_2", all.x=T, sort=F)

mouse_to_human=a1
mouse_to_human = subset(mouse_to_human, !is.na(GeneSymbol) & !is.na(symbol_1) & !symbol_1 == "")
rownames(mouse_to_human) = mouse_to_human$GeneSymbol
```


# Plots for selected gene sets

```{r}
selected_genesets = c("Lgr5_SC_Signature_Munoz", "Yui_DSS_colitis_vs_control_down", "Yui_DSS_colitis_vs_control_up", "Gregorieff_Yap_KO_up", "Gregorieff_Yap_KO_down", "Gregorieff_Yap_Tg_up", "Gregorieff_Yap_Tg_down")
```

```{r, fig.width=16, fig.height=6, results="hide"}
all_plots = list()
for (sel_comparison in names(all_results)) {
  tmp = all_results[[sel_comparison]]
  # tmp2 = merge(tmp, mouse_to_human, by="GeneSymbol")
  # tmp2 = subset(tmp2, !is.na(symbol_1))
  ranks_mouse = unlist(tapply(tmp$t, tmp$GeneSymbol, function(x) x[which(abs(x)==max(abs(x)))] ))

  tmp2 = merge(subset(tmp, !is.na(GeneSymbol)), mouse_to_human, by="GeneSymbol")
  tmp2 = subset(tmp2, !is.na(symbol_1))
  ranks_human = unlist(tapply(tmp2$t, tmp2$symbol_1, function(x) x[which(abs(x)==max(abs(x)))] ))

  gg <- strsplit(sel_comparison, "_vs_") |> unlist()
  up_group <- gg[1]
  down_group <- gg[2]
  
  for (g in selected_genesets) {
    rr = subset(all_gsea_results[[sel_comparison]], pathway==g)
    loc = unlist(lapply(gset_collections_human, function(x) grep(paste0("^",g,"$"), names(x) )))
    if(length(loc)>0) {
      sel_genes = unlist(gset_collections_human[[names(loc)]][loc])
      ranks = ranks_human
    } else {
      loc = unlist(lapply(gset_collections_mouse, function(x) grep(paste0("^",g,"$"), names(x) )))
      sel_genes = unlist(gset_collections_mouse[[names(loc)]][loc])
      ranks = ranks_mouse

    }
    p = myPlotEnrichment(sel_genes, ranks, group_pos = up_group, group_neg = down_group)
    p1 = p + ggtitle(paste0(g, "\n", sel_comparison), paste("FDR=",prettyNum(rr$FDR_global, digits=2),", p=",prettyNum(rr$pval, digits=2), ", ES=",prettyNum(rr$ES, digits=2), ", NES=",prettyNum(rr$NES, digits=2) ) )  + theme(title = element_text(size=12))
    #print(p1)
    all_plots[[paste0(g,"_",sel_comparison)]] = p1
  }  
}

all_plots = all_plots[order(names(all_plots))]
print(marrangeGrob(all_plots, ncol=3, nrow=1))

```


### PDF export

All plots have been exported to corresponding files in PDF format in folder `r result_folder`.

```{r}
for (sel_comparison in names(all_results)) {
  for (g in selected_genesets) {
    tmp = all_results[[sel_comparison]]

    ranks_mouse = unlist(tapply(tmp$t, tmp$GeneSymbol, function(x) x[which(abs(x)==max(abs(x)))] ))

    tmp2 = merge(subset(tmp, !is.na(GeneSymbol)), mouse_to_human, by="GeneSymbol")
    tmp2 = subset(tmp2, !is.na(symbol_1))
    ranks_human = unlist(tapply(tmp2$t, tmp2$symbol_1, function(x) x[which(abs(x)==max(abs(x)))] ))
    
    
    gg <- strsplit(sel_comparison, "_vs_") |> unlist()
    up_group <- gg[1]
    down_group <- gg[2]
    
    
    rr = subset(all_gsea_results[[sel_comparison]], pathway==g)
      loc = unlist(lapply(gset_collections_human, function(x) grep(paste0("^",g,"$"), names(x) )))
      if(length(loc)>0) {
        sel_genes = unlist(gset_collections_human[[names(loc)]][loc])
        ranks = ranks_human
      } else {
        loc = unlist(lapply(gset_collections_mouse, function(x) grep(paste0("^",g,"$"), names(x) )))
        sel_genes = unlist(gset_collections_mouse[[names(loc)]][loc])
        ranks = ranks_mouse
  
    }
    p = myPlotEnrichment(sel_genes, ranks, linewidth = 2, group_pos = up_group, group_neg = down_group)
    pdf(file=file.path(result_folder, paste0(sel_comparison,"_",g, ".pdf")), width=8, height = 6)
    print(p + ggtitle(g, paste("FDR=",prettyNum(rr$FDR_global, digits=2),", p=",prettyNum(rr$pval, digits=2), ", ES=",prettyNum(rr$ES, digits=2), ", NES=",prettyNum(rr$NES, digits=2) ) ) )
    #print(p + ggtitle(g, paste("Adj.pval=",prettyNum(rr$padj, digits=2), "ES=",prettyNum(rr$ES, digits=2), ", NES=",prettyNum(rr$NES, digits=2) ) ) )
    dev.off()
  }
}  
```



# Software versions

```{r}
sessionInfo()
```


