---
title: "Microarray Data of P53KO/KOWI - GSEA analysis on DGE results - Heatmap"
author: "Hilmar Berger"
output: 
 html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    number_sections: true
    code_folding: hide
    
pdf_document:
    fig_caption: true
date: "`r format(Sys.time(), '%d %B, %Y %H:%M:%S')`"
---

```{r, warning=FALSE, message=FALSE}
rm(list=ls())

#suppressMessages(library(xlsx))
library(ggplot2)
library(knitr)
library(scales)
library(fgsea)
library(data.table)

library(RColorBrewer)
library(pheatmap)

result_folder = "../../Results/Organoids_Trp53_KO_KOWI"
if (!file.exists(result_folder)) dir.create(result_folder, recursive=T)

set.seed(12345)
```

# Overview

This is data from a stomach antrum samples of various mouse experiments hybridized as dual channel libraries to Agilent 014568 micro arrays. 

For analysis, a preranked analysis using the R package fGSEA was used that should give similar results to preranked analysis in standard GSEA. 



```{r}
#data_generation_date = "2022-01-13"
output_folder = result_folder
GSEA_datafile = file.path(output_folder,"fGSEA_Results.Rdata")
load(GSEA_datafile)
load(file.path(output_folder, "Geneset_collections.Rdata"))
```



```{r}
sel_genesets_1 = "GAO_LARGE_INTESTINE_24W_C10_ENTEROCYTE
GAO_LARGE_INTESTINE_24W_C8_GOBLET_CELL
GAO_LARGE_INTESTINE_24W_C5_LGR5POS_STEM_CELL
REACTOME_PYRUVATE_METABOLISM_AND_CITRIC_ACID_TCA_CYCLE
HALLMARK_OXIDATIVE_PHOSPHORYLATION"
selected_genesets =  c(unlist(strsplit(sel_genesets_1, "\n")) , names(gset_collections_mouse[["CUSTOM_MOUSE"]]), names(gset_collections_human[["CUSTOM_HUMAN"]]) )
```


```{r}
results_combined = do.call(rbind, Map(function(n) {tmp = all_gsea_results[[n]]; tmp$comparison=n; return(tmp)}, names(all_gsea_results)))
```

```{r}
selected_comparisons = names(all_gsea_results)
```


# All NES values

```{r, fig.height=7}
tmp = subset(results_combined, comparison %in% selected_comparisons & pathway %in% selected_genesets)

tmp2 = reshape2::dcast(tmp, pathway ~ comparison, value.var="NES")
gsea_mat = as.matrix(tmp2[, 2:ncol(tmp2)])
rownames(gsea_mat) <- tmp2$pathway

pheatmap(gsea_mat)
```

# only significant NES values, setting non-significant to 0

```{r, fig.height=8}
tmp = subset(results_combined, comparison %in% selected_comparisons & pathway %in% selected_genesets & FDR_global < 0.1)

tmp2 = reshape2::dcast(tmp, pathway ~ comparison, value.var="NES")
gsea_mat = as.matrix(tmp2[, 2:ncol(tmp2)])
rownames(gsea_mat) <- tmp2$pathway

gsea_mat[is.na(gsea_mat)] <- 0
cp = colorRampPalette(c("red","grey90","blue"))
pheatmap(gsea_mat, color =  cp(100))
```

# Software versions

```{r}
sessionInfo()
```
